Combinatorial Proofs and Decomposition Theorems for First-order Logic

### 2021-01-18 main.tex
First comments on IX.B

\begin{proof}[Proof of Lemma~\ref{skew->cw}]
  By Lemma~\ref{lem:PE} we have $\gA=\graphof{\PE A}$ and
  $\gB=\graphof{\PE B}$.  Let $\vB'\subseteq\vB$ be the image of
  $\phi$, and let $\gB_1$ be the subgraph of $\gB$ induced by
  $\vB'$. Hence, we have two maps $\phi''\colon\gA\to\gB_1$ being a
  surjection and $\phi'\colon \gB_1\to\gB$ being an injection that
  reflects edges. Both, $\phi'$ and $\phi''$ remain skew
  bifibrations. Let us first look at $\phi'$.  Let $\PE B_1$ be the
  propositional formula obtained from $\PE B$ by removing all atoms
  that are not represented by vertices in $\vB'$. Then $\graphof{\PE
    B_1}=\gB_1$.  By \cite[Proposition~7.6.1]{str:07:RTA}, we have a
  derivation $\vlder{\set{\wrD,\fequ}}{\PE\Deri_1}{\PE B}{\PE B_1}$.
  A subformula of $\PE B$ is called \emph{weak} if it has been
  introduced by a weakening. All subformulas of a weak formula are
  also weak. Two weak subformulas $B'$ and $B''$ of $\PE B$ form a
  \emph{weak pair} if $B\fequ\Scons{B'\vlor B''}$ for some context
  $\Sconhole$. We now can assume without loss of generality that
  whenever weak subformulas $B'$ and $B''$ for a weak pair, they have
  been introduced by the same instance of $\wlD$ or $\wrD$ in
  $\PE\Deri_1$.\footnote{If $\PE\Deri_1$ is not of that shape, it can
  brought into this form by simple rule permutations, and then
  collapsing several weakenings into a single one.}  Now we show that
  $\PE\Deri_1$ can be lifted. For this, observe that whenever a
  weakening in $\PE\Deri_1$ deletes an atom $x\in\VAR$, it must also
  delete all atoms in the scope of the corresponding quantifier,
  because $\phi$ is a fibration on the binding graph. Hence, each line
  in $\PE\Deri_1$ is the propositional encoding $\PE P$ of a
  first-order formula $P$.  We now have to show
  that each instance of $\wrD$ is indeed a correct
  application in first-order logic. For this we have to inspect the
  cases a weakening happens inside a subformula $x\vlor C$ or $x\vlan
  C$ in $\PE\Deri_1$. There are the following cases:
  \begin{equation*}%\small
    \vlinf{}{}{S\cons{x\vlor D\vlor C}}{S\cons{x\vlor C}}
    \qquad
    \vlinf{}{}{S\cons{x\vlan (D\vlor C)}}{S\cons{x\vlan C}}
    \qquad
    \vlinf{}{}{S\cons{(x\vlor D)\vlan C}}{S\cons{x\vlan C}}
  \end{equation*}
  In the first case the weakening happens inside the scope of a
  $\forall$-quantifier, and in the second case inside the scope of a
  $\exists$-quantifier. Both cases are unproblematic in first-order
  logic. However, in the third case an $\exists$-quantifier would be
  transformed into an $\forall$-quantifier. But as $\phi$ has to
  preserve existentials, this third case cannot occur.
  Thus we have a first order derivation
  $\vlder{\set{\wrD,\fequ}}{\Deri_1}{B}{B_1}$.

@@

  inferene rule

>>>>

\begin{proof}[Proof of Lemma~\ref{skew->cw}]
  By Lemma~\ref{lem:PE} we have $\gA=\graphof{\PE A}$ and
  $\gB=\graphof{\PE B}$.  Let $\vB'\subseteq\vB$ be the image of
  $\phi$, and let $\gB_1$ be the subgraph of $\gB$ induced by
  $\vB'$. Hence, we have two maps $\phi''\colon\gA\to\gB_1$ being a
  surjection and $\phi'\colon \gB_1\to\gB$ being an injection that
  reflects edges. \juihsuan{what do you mean by "reflect edges"?} Both, $\phi'$ and $\phi''$ remain skew
  bifibrations. Let us first look at $\phi'$.  Let $\PE B_1$ be the
  propositional formula obtained from $\PE B$ by removing all atoms
  that are not represented by vertices in $\vB'$. Then $\graphof{\PE
    B_1}=\gB_1$.  By \cite[Proposition~7.6.1]{str:07:RTA}, we have a
  derivation $\vlder{\set{\wrD,\fequ}}{\PE\Deri_1}{\PE B}{\PE B_1}$.
  A subformula of $\PE B$ is called \emph{weak} if it has been
  introduced by a weakening. All subformulas of a weak formula are
  also weak. Two weak subformulas $B'$ and $B''$ of $\PE B$ form a
  \emph{weak pair} if $B\fequ\Scons{B'\vlor B''}$ for some context
  $\Sconhole$. We now can assume without loss of generality that
  whenever weak subformulas $B'$ and $B''$ for a weak pair, they have
  been introduced by the same instance of $\wrD$ in
  $\PE\Deri_1$.\footnote{If $\PE\Deri_1$ is not of that shape, it can
  brought into this form by simple rule permutations, and then
  collapsing several weakenings into a single one.}  Now we show that
  $\PE\Deri_1$ can be lifted. For this, observe that whenever a
  weakening in $\PE\Deri_1$ deletes an atom $x\in\VAR$, it must also
  delete all atoms in the scope of the corresponding quantifier,
  because $\phi$ \juihsuan{should be $\phi'$ here} is a fibration on the binding graph. Hence, each line
  in $\PE\Deri_1$ is the propositional encoding $\PE P$ of a
  first-order formula $P$.  We now have to show
  that each instance of $\wrD$ is indeed a correct
  application in first-order logic. For this we have to inspect the
  cases a weakening happens inside a subformula $x\vlor C$ or $x\vlan
  C$ in $\PE\Deri_1$. There are the following cases:
  \begin{equation*}%\small
    \vlinf{}{}{S\cons{x\vlor D\vlor C}}{S\cons{x\vlor C}}
    \qquad
    \vlinf{}{}{S\cons{x\vlan (D\vlor C)}}{S\cons{x\vlan C}}
    \qquad
    \vlinf{}{}{S\cons{(x\vlor D)\vlan C}}{S\cons{x\vlan C}}
  In the first case the weakening happens inside the scope of a
  $\forall$-quantifier, and in the second case inside the scope of a
  $\exists$-quantifier. Both cases are unproblematic in first-order
  logic. However, in the third case an $\exists$-quantifier would be
  transformed into an $\forall$-quantifier. But as $\phi$ has to
  preserve existentials, this third case cannot occur.
  Thus we have a first order derivation
  $\vlder{\set{\wrD,\fequ}}{\Deri_1}{B}{B_1}$.
  \juihsuan{$B_1$ is not defined (The notation $\PE B_1$ suggests that it is the
propositional encoding of some formula but we need to prove it). This can be
proven using the fact that $\phi'$ is a skew bifibration here.}

@@

  inference rule
